---
title: 'RepResearch: Course Project 2 - Storm Data Analysis'
output: html_document
---

## R Markdown

This is my submission for the 2nd course project for the Reproducible Research course in Coursera. In this project, I have analysed the storm data to infer fatalities, injuries and economic damage.

This first chunk covers the data extraction, manipulation and plot creation part of the exercise. The subsequent chunks will present the plots created in this code chunk.

```{r Data_Extraction, echo=TRUE}

library(ggplot2)
library(gridExtra)
stormdata <- read.csv("/Users/wolfram/Downloads/repdata-data-StormData.csv", header = TRUE, sep = ',')

df <- data.frame(stormdata)

fatalities <- aggregate(df$FATALITIES,by = list(df$EVTYPE), FUN = sum)
injuries <- aggregate(df$INJURIES, by = list(df$EVTYPE), FUN = sum)

cropdmg <- aggregate(df$CROPDMG, by = list(df$EVTYPE), FUN = sum)
prpdmg <- aggregate(df$PROPDMG, by = list(df$EVTYPE), FUN = sum)
ecodmg <- aggregate(df$CROPDMG+df$PROPDMG, by = list(df$EVTYPE), FUN = sum)

fat_df <- data.frame(fatalities)
inj_df <- data.frame(injuries)
eco_df <- data.frame(ecodmg)

colnames(fatalities) <- c("evtype","sum")
colnames(injuries) <- c("evtype", "sum")
colnames(eco_df) <- c("evtype", "sum")

fatalities <- fatalities[with(fatalities, order(-sum)),]
injuries <- injuries[with(injuries, order(-sum)),]
ecodamage <- eco_df[with(eco_df, order(-sum)),]

top_fat <- fatalities[1:20,]
top_inj <- injuries[1:20,]

```

Plot for Injuries and Fatalities caused by the Top 20 Event Types

```{r Plot_1, echo= TRUE, fig.height= 20, fig.width= 30}
fats <- ggplot(top_fat, aes(evtype,y = count, color = variable)) + geom_point(aes(y = sum, col = "sum")) + theme(axis.text.x=element_text(angle=60, hjust =1))+ggtitle("Top Fatality Factors in the Country") 

injs <- ggplot(top_inj, aes(evtype,y = count, color = variable)) + geom_point(aes(y = sum, col = "sum")) + theme(axis.text.x=element_text(angle=60, hjust =1))+ggtitle("Top Injury Factors in the Country")

grid.arrange(fats, injs, ncol = 1)
```

Plot for Economic Damages caused by the top 20 Event Types

``` {r Plot_2, echo= TRUE}
ecods <- ggplot(ecodamage[1:20,], aes(evtype, y = Amount, color = variable)) + geom_point(aes(y = sum/1000000, col = "sum")) + theme(axis.text.x=element_text(angle=60, hjust =1))+ggtitle("Top Economic Damage Contributors")

ecods
```
